---
title: "3 - Modeling"
author: "Micah Jona"
date: "2/3/2021"
output: html_document
---

To create a quality of contact metric, I'm creating a multi-class classification model that aims to predict the outcome of each batted ball using a XGBoost model. XGBoost is a decision tree based ensemble supervised machine learning algorithm that uses a gradient boosting framework. Gradient boosting is a case of boosting where errors are minimized by gradient descent and applies the principle of boosting weak learners. The rationale for using XGBoost lies in its easy to understand nature and the structure and size of the data. Decision trees for classifications are easy to interpret on a fundamental level. A basic decision tree, for example just a tree stump, visualized it is easy to see the rationale for the classification. Since this dataset is relatively large (over 100,000 observations per season) and is tabular data, gradient boosting machines (GBMs), specifically XGBoost seems like a fine choice for this problem. Also, GBMs are non-parametric models that are suitable for this non-linear data. 

For my modeling, I'm going to use play outcome as the response variable. The play outcome (out, single, double, triple, homerun) is a good starting place for measuring quality of contact. If we are just looking at grading quality of contact based on the result of the batted ball then we know that a homerun is better than a triple which is better than a double and so on. This metric provides a broadly accepted way to rank the quality of a batted ball. Generally speaking, hitting a ball harder (more exit speed) is better than hitting a ball softer. However, it’s important to include the launch angle when using exit speed. A batter would much rather hit a ball 105 mph at 25 degrees than a ball 125 mph at -10 degrees. There are exceptions, but in the aggregate we know that homeruns have the best combination of exit speed and launch angle. However, there is more to assessing quality of contact than just exit speed and launch angle, such as launch direction, bearing, and distance. Each of the metrics individually can provide some insight about the quality of contact, but when taken together it quickly becomes intractable to rank every possible combination. From this, I felt that all of these metrics were best captured and ranked by the play outcome.

I start the modeling by doing a simple 10-fold cross-validation to determine the appropriate number of iterations to use in a baseline XGBoost model. There are two main choices I made before starting the XGBoost modeling. I'm going to use multi:softprob as the objective of the model with mlogloss evaluation metric. multi:softprob returns the probability of each observation being a member of each class. There are five classes of the response variable (out, single, double, triple, and homerun). I made this choice because you can gain more information about each batted ball knowing the probability of it belonging to each class rather than the model returning which class has the maximum probability. The evaluation metric is used for validation data in both the cross-validation and XGBoost modeling for both early stopping and model evaluation. mlogloss is the loss function used in multinomial logistic regression and is defined as the negative log-likelihood of a logistic model that returns probabilities for its training data. The baseline model also has no hyperparameters specified, meaning that all of those parameters are their default values. Normally, I'd hyperparameter tune an XGBoost model using a latin hypercube, however I don't want to subject my laptop to running 150 different models and in my experience, tuning seems to increase accuracy by at most 1%. 

Even though I'm going to run a model for each season on the entire season's worth of data, I'm still going to do a typical 70/30 train/test split first to ensure the model is not over- or under-fitting. 

In addition, to the XGBoost model, I'm going use the class probabilties (probability of out, single, double, triple, and homerun) as inputs to a generalized linear mixed-effects model. The multinomial modeling workflow comes from *Baseball Prospectus* and their creation of DRC+ (https://www.baseballprospectus.com/news/article/48293/entirely-beyond-wowy-a-breakdown-of-drc/#fnref4). Below I'm going to create four binomial subsets, therefore turning this multinomial problem into multiple binomial models. The four susbets are: out + single, out + double, out + triple, and out + homerun. Each subset has outs because that is our "pivot," the common category every other event will be regressed against. Outs were selected because, as Judge says, this methodology works best when the pivot is also the largest category. The most common outcome for a batted ball is an out (remember the typical league average batting average of balls in play (BABIP) is ~.300). Then each binomial subset is fit on a generalized linear mixed-effects model. My singular motivation of using this modeling (besides saying I'm using a generalized linear mixed-effects model) is to include batter and pitcher identity. It's clear that batter and pitchers have some effect on the outcome of a batted ball. For example, even without explicitly including sprint speed, in my experience this type of model is able to pick up on what batters are able to turn singles into double or doubles into triples. It's theoretically possible to include batter and pitchers in my XGBoost modeling as dummy variables, but I don't think my poor laptop would've been able to handle a 105,000 by 3,200 dataset five times for gradient boosting. The final equation for the model is: outcome ~ (1|pitcher) + (1|batter) + class.probability. The class.prob is dependent on what binomial subset the model is running on. In the single subset that class.probability would be single.probability, for doubles it would be double.probability, etc. In addition to including batter and pitcher identity, which I believe will make the model a lot more robust, this model also acts as a sort of stacking algorithm.

Stacking is a popular ensemble method in machine learning. Generally, stacking involves training several models with different algorithms that are relatively simple (base-learners). Then the predictions from each model are fed into one final model (meta-learner) to make the final prediction. So the inputs for the meta-learner are the prediction outputs of the base-learners. The benefit of stacking is that can combine the capabilities of a multitude of good models that end up making a stronger model than any of the base-learners. I wouldn't consider my methodology as exactly stacking because I'm not combining models, rather using the outputs from one model, in addition to other features, as the input to a final model. However, I do think that I still enjoy some of the benefits of stacking. It's important to note, as it was pointed out to me, that sometimes XGBoost can smooth over the linear model signal for the random effects (pitcher and batter identity). Therefore, it's a possibility that the mixed-effects model doesn't entirely capture the random effect of each batter and pitcher accurrately. 


```{r message=FALSE, warning=FALSE}
library(baseballr)
library(magrittr)
library(tidyverse)
library(ggplot2)
library(ggthemes)
library(patchwork)
library(arm)
library(blme)
library(lme4)
library(xgboost)
library(ggrepel)
library(fastDummies)
library(caret)
```


```{r}
batted_ball_model <- read_csv("batted_ball_processed.csv") %>% dplyr::select(-c("X1"))
```


```{r}
# setting seed for reproducibility
set.seed(123)
```

Creating a dataset for each season.
```{r}
batted_ball_model_2015 <- batted_ball_model %>% filter(game_year == 2015)
batted_ball_model_2016 <- batted_ball_model %>% filter(game_year == 2016)
batted_ball_model_2017 <- batted_ball_model %>% filter(game_year == 2017)
batted_ball_model_2018 <- batted_ball_model %>% filter(game_year == 2018)
batted_ball_model_2019 <- batted_ball_model %>% filter(game_year == 2019)
```

Here is the XGBoost modeling for each season, split into train/test sets to determine overall generalization ability. I'm only going to show the 2015 model, as the rest are the same.
```{r}
####### 2015
# create indecies for splitting (need to use a specific column from df for it to run)
train_index_2015 <- createDataPartition(y = batted_ball_model_2015$outcome_ec, p = .70, list = FALSE) %>% as.vector()

# split into train and test
train_2015 <- batted_ball_model_2015[train_index_2015,]
test_2015 <- batted_ball_model_2015[-train_index_2015,]

# need a vector of response variable (outcome)
vec_label_train_2015 <- train_2015$outcome_ec

train_data_2015 <- as.matrix(train_2015 %>% dplyr::select(c("launch_angle", "launch_speed", "spray_angle", "stand_r", "p_throws_r", "temperature", "is_dome",
                                                            pitch_type_CH:venue_batter.stand_YankeeStadium_r)))

vec_label_test_2015 <- test_2015$outcome_ec
test_data_2015 <- as.matrix(test_2015 %>% dplyr::select(c("launch_angle", "launch_speed", "spray_angle", "stand_r", "p_throws_r", "temperature", "is_dome",
                                                            pitch_type_CH:venue_batter.stand_YankeeStadium_r)))

# convert the train and test into xgb.DMatrix
x_train_2015 = xgboost::xgb.DMatrix(data = train_data_2015, label = vec_label_train_2015)
x_test_2015 = xgboost::xgb.DMatrix(data = test_data_2015, label = vec_label_test_2015)

baseline_cv_2015 <- xgboost::xgb.cv(params = list(objective = "multi:softprob", eval_metric = c("mlogloss"), num_class = 5), data = x_train_2015, nrounds = 2500, early_stopping_rounds = 10, nfold=10, stratified = T)

base_mod_2015 <- xgboost::xgboost(params = list(objective = "multi:softprob", eval_metric = c("mlogloss"), num_class = 5), data = x_train_2015, nrounds = baseline_cv_2015$best_ntreelimit)
```

```{r include=FALSE}
####### 2016

# create indecies for splitting (need to use a specific column from df for it to run)
train_index_2016 <- createDataPartition(y = batted_ball_model_2016$outcome_ec, p = .70, list = FALSE) %>% as.vector()

# split into train and test
train_2016 <- batted_ball_model_2016[train_index_2016,]
test_2016 <- batted_ball_model_2016[-train_index_2016,]

# need a vector of response variable (outcome)
vec_label_train_2016 <- train_2016$outcome_ec

train_data_2016 <- as.matrix(train_2016 %>% dplyr::select(c("launch_angle", "launch_speed", "spray_angle", "stand_r", "p_throws_r", "temperature", "is_dome",
                                                            pitch_type_CH:venue_batter.stand_YankeeStadium_r)))

vec_label_test_2016 <- test_2016$outcome_ec
test_data_2016 <- as.matrix(test_2016 %>% dplyr::select(c("launch_angle", "launch_speed", "spray_angle", "stand_r", "p_throws_r", "temperature", "is_dome",
                                                            pitch_type_CH:venue_batter.stand_YankeeStadium_r)))

# convert the train and test into xgb.DMatrix
x_train_2016 = xgboost::xgb.DMatrix(data = train_data_2016, label = vec_label_train_2016)
x_test_2016 = xgboost::xgb.DMatrix(data = test_data_2016, label = vec_label_test_2016)

baseline_cv_2016 <- xgboost::xgb.cv(params = list(objective = "multi:softprob", eval_metric = c("mlogloss"), num_class = 5), data = x_train_2016, nrounds = 2500, early_stopping_rounds = 10, nfold=10, stratified = T)

base_mod_2016 <- xgboost::xgboost(params = list(objective = "multi:softprob", eval_metric = c("mlogloss"), num_class = 5), data = x_train_2016, nrounds = baseline_cv_2016$best_ntreelimit)

####### 2017

# create indecies for splitting (need to use a specific column from df for it to run)
train_index_2017 <- createDataPartition(y = batted_ball_model_2017$outcome_ec, p = .70, list = FALSE) %>% as.vector()

# split into train and test
train_2017 <- batted_ball_model_2017[train_index_2017,]
test_2017 <- batted_ball_model_2017[-train_index_2017,]

# need a vector of response variable (outcome)
vec_label_train_2017 <- train_2017$outcome_ec

train_data_2017 <- as.matrix(train_2017 %>% dplyr::select(c("launch_angle", "launch_speed", "spray_angle", "stand_r", "p_throws_r", "temperature", "is_dome",
                                                            pitch_type_CH:venue_batter.stand_YankeeStadium_r)))

vec_label_test_2017 <- test_2017$outcome_ec
test_data_2017 <- as.matrix(test_2017 %>% dplyr::select(c("launch_angle", "launch_speed", "spray_angle", "stand_r", "p_throws_r", "temperature", "is_dome",
                                                            pitch_type_CH:venue_batter.stand_YankeeStadium_r)))

# convert the train and test into xgb.DMatrix
x_train_2017 = xgboost::xgb.DMatrix(data = train_data_2017, label = vec_label_train_2017)
x_test_2017 = xgboost::xgb.DMatrix(data = test_data_2017, label = vec_label_test_2017)

baseline_cv_2017 <- xgboost::xgb.cv(params = list(objective = "multi:softprob", eval_metric = c("mlogloss"), num_class = 5), data = x_train_2017, nrounds = 2500, early_stopping_rounds = 10, nfold=10, stratified = T)

base_mod_2017 <- xgboost::xgboost(params = list(objective = "multi:softprob", eval_metric = c("mlogloss"), num_class = 5), data = x_train_2017, nrounds = baseline_cv_2017$best_ntreelimit)

####### 2018

# create indecies for splitting (need to use a specific column from df for it to run)
train_index_2018 <- createDataPartition(y = batted_ball_model_2018$outcome_ec, p = .70, list = FALSE) %>% as.vector()

# split into train and test
train_2018 <- batted_ball_model_2018[train_index_2018,]
test_2018 <- batted_ball_model_2018[-train_index_2018,]

# need a vector of response variable (outcome)
vec_label_train_2018 <- train_2018$outcome_ec

train_data_2018 <- as.matrix(train_2018 %>% dplyr::select(c("launch_angle", "launch_speed", "spray_angle", "stand_r", "p_throws_r", "temperature", "is_dome",
                                                            pitch_type_CH:venue_batter.stand_YankeeStadium_r)))

vec_label_test_2018 <- test_2018$outcome_ec
test_data_2018 <- as.matrix(test_2018 %>% dplyr::select(c("launch_angle", "launch_speed", "spray_angle", "stand_r", "p_throws_r", "temperature", "is_dome",
                                                            pitch_type_CH:venue_batter.stand_YankeeStadium_r)))

# convert the train and test into xgb.DMatrix
x_train_2018 = xgboost::xgb.DMatrix(data = train_data_2018, label = vec_label_train_2018)
x_test_2018 = xgboost::xgb.DMatrix(data = test_data_2018, label = vec_label_test_2018)

baseline_cv_2018 <- xgboost::xgb.cv(params = list(objective = "multi:softprob", eval_metric = c("mlogloss"), num_class = 5), data = x_train_2018, nrounds = 2500, early_stopping_rounds = 10, nfold=10, stratified = T)

base_mod_2018 <- xgboost::xgboost(params = list(objective = "multi:softprob", eval_metric = c("mlogloss"), num_class = 5), data = x_train_2018, nrounds = baseline_cv_2018$best_ntreelimit)

####### 2019

# create indecies for splitting (need to use a specific column from df for it to run)
train_index_2019 <- createDataPartition(y = batted_ball_model_2019$outcome_ec, p = .70, list = FALSE) %>% as.vector()

# split into train and test
train_2019 <- batted_ball_model_2019[train_index_2019,]
test_2019 <- batted_ball_model_2019[-train_index_2019,]

# need a vector of response variable (outcome)
vec_label_train_2019 <- train_2019$outcome_ec

train_data_2019 <- as.matrix(train_2019 %>% dplyr::select(c("launch_angle", "launch_speed", "spray_angle", "stand_r", "p_throws_r", "temperature", "is_dome",
                                                            pitch_type_CH:venue_batter.stand_YankeeStadium_r)))

vec_label_test_2019 <- test_2019$outcome_ec
test_data_2019 <- as.matrix(test_2019 %>% dplyr::select(c("launch_angle", "launch_speed", "spray_angle", "stand_r", "p_throws_r", "temperature", "is_dome",
                                                            pitch_type_CH:venue_batter.stand_YankeeStadium_r)))

# convert the train and test into xgb.DMatrix
x_train_2019 = xgboost::xgb.DMatrix(data = train_data_2019, label = vec_label_train_2019)
x_test_2019 = xgboost::xgb.DMatrix(data = test_data_2019, label = vec_label_test_2019)

baseline_cv_2019 <- xgboost::xgb.cv(params = list(objective = "multi:softprob", eval_metric = c("mlogloss"), num_class = 5), data = x_train_2019, nrounds = 2500, early_stopping_rounds = 10, nfold=10, stratified = T)

base_mod_2019 <- xgboost::xgboost(params = list(objective = "multi:softprob", eval_metric = c("mlogloss"), num_class = 5), data = x_train_2019, nrounds = baseline_cv_2019$best_ntreelimit)
```


Here I'm computing the test set accuracy for each season's model. The accuracies range from 80.8% to 81.6%, which is very good. Keep in mind that a naive classifier would have an accuracy around 20% (randomly guessing the class for each observation). 
```{r}
xgb.pred_2015 = predict(base_mod_2015, x_test_2015, reshape=T)
xgb.pred_2015 = as.data.frame(xgb.pred_2015)
colnames(xgb.pred_2015) = c("Out", "Single", "Double", "Triple", "Home_Run")

preds_base_2015 <- test_2015
preds_base_2015$out.prob <- xgb.pred_2015[,1]
preds_base_2015$single.prob <- xgb.pred_2015[,2]
preds_base_2015$double.prob <- xgb.pred_2015[,3]
preds_base_2015$triple.prob <- xgb.pred_2015[,4]
preds_base_2015$hr.prob <- xgb.pred_2015[,5]

max_prob_2015 <- colnames(xgb.pred_2015)[apply(xgb.pred_2015,1,which.max)]

preds_base_2015$max.prob <- max_prob_2015

preds_base_2015$pred.right <- ifelse(preds_base_2015$max.prob == "Out" & preds_base_2015$outcome_ec == 0, "yes", 
                                ifelse(preds_base_2015$max.prob == "Single" & preds_base_2015$outcome_ec == 1, "yes",
                                       ifelse(preds_base_2015$max.prob == "Double" & preds_base_2015$outcome_ec == 2, "yes",
                                              ifelse(preds_base_2015$max.prob == "Triple" & preds_base_2015$outcome_ec == 3, "yes",
                                                     ifelse(preds_base_2015$max.prob == "Home_Run" & preds_base_2015$outcome_ec == 4, "yes", "no")))))
preds_base_2015$pred.right <- as.factor(preds_base_2015$pred.right)

sum(preds_base_2015$pred.right == "yes") / nrow(preds_base_2015) #0.8083159

preds_base_2015 %>%
  group_by(outcome_ec, pred.right) %>%
  count()
```

```{r include=FALSE}
###### 2016
xgb.pred_2016 = predict(base_mod_2016, x_test_2016, reshape=T)
xgb.pred_2016 = as.data.frame(xgb.pred_2016)
colnames(xgb.pred_2016) = c("Out", "Single", "Double", "Triple", "Home_Run")

preds_base_2016 <- test_2016
preds_base_2016$out.prob <- xgb.pred_2016[,1]
preds_base_2016$single.prob <- xgb.pred_2016[,2]
preds_base_2016$double.prob <- xgb.pred_2016[,3]
preds_base_2016$triple.prob <- xgb.pred_2016[,4]
preds_base_2016$hr.prob <- xgb.pred_2016[,5]

max_prob_2016 <- colnames(xgb.pred_2016)[apply(xgb.pred_2016,1,which.max)]

preds_base_2016$max.prob <- max_prob_2016

preds_base_2016$pred.right <- ifelse(preds_base_2016$max.prob == "Out" & preds_base_2016$outcome_ec == 0, "yes", 
                                ifelse(preds_base_2016$max.prob == "Single" & preds_base_2016$outcome_ec == 1, "yes",
                                       ifelse(preds_base_2016$max.prob == "Double" & preds_base_2016$outcome_ec == 2, "yes",
                                              ifelse(preds_base_2016$max.prob == "Triple" & preds_base_2016$outcome_ec == 3, "yes",
                                                     ifelse(preds_base_2016$max.prob == "Home_Run" & preds_base_2016$outcome_ec == 4, "yes", "no")))))
preds_base_2016$pred.right <- as.factor(preds_base_2016$pred.right)

sum(preds_base_2016$pred.right == "yes") / nrow(preds_base_2016) #0.8102002

preds_base_2016 %>%
  group_by(outcome_ec, pred.right) %>%
  count()

###### 2017
xgb.pred_2017 = predict(base_mod_2017, x_test_2017, reshape=T)
xgb.pred_2017 = as.data.frame(xgb.pred_2017)
colnames(xgb.pred_2017) = c("Out", "Single", "Double", "Triple", "Home_Run")

preds_base_2017 <- test_2017
preds_base_2017$out.prob <- xgb.pred_2017[,1]
preds_base_2017$single.prob <- xgb.pred_2017[,2]
preds_base_2017$double.prob <- xgb.pred_2017[,3]
preds_base_2017$triple.prob <- xgb.pred_2017[,4]
preds_base_2017$hr.prob <- xgb.pred_2017[,5]

max_prob_2017 <- colnames(xgb.pred_2017)[apply(xgb.pred_2017,1,which.max)]

preds_base_2017$max.prob <- max_prob_2017

preds_base_2017$pred.right <- ifelse(preds_base_2017$max.prob == "Out" & preds_base_2017$outcome_ec == 0, "yes", 
                                ifelse(preds_base_2017$max.prob == "Single" & preds_base_2017$outcome_ec == 1, "yes",
                                       ifelse(preds_base_2017$max.prob == "Double" & preds_base_2017$outcome_ec == 2, "yes",
                                              ifelse(preds_base_2017$max.prob == "Triple" & preds_base_2017$outcome_ec == 3, "yes",
                                                     ifelse(preds_base_2017$max.prob == "Home_Run" & preds_base_2017$outcome_ec == 4, "yes", "no")))))
preds_base_2017$pred.right <- as.factor(preds_base_2017$pred.right)

sum(preds_base_2017$pred.right == "yes") / nrow(preds_base_2017) #0.8154189

preds_base_2017 %>%
  group_by(outcome_ec, pred.right) %>%
  count()

###### 2018
xgb.pred_2018 = predict(base_mod_2018, x_test_2018, reshape=T)
xgb.pred_2018 = as.data.frame(xgb.pred_2018)
colnames(xgb.pred_2018) = c("Out", "Single", "Double", "Triple", "Home_Run")

preds_base_2018 <- test_2018
preds_base_2018$out.prob <- xgb.pred_2018[,1]
preds_base_2018$single.prob <- xgb.pred_2018[,2]
preds_base_2018$double.prob <- xgb.pred_2018[,3]
preds_base_2018$triple.prob <- xgb.pred_2018[,4]
preds_base_2018$hr.prob <- xgb.pred_2018[,5]

max_prob_2018 <- colnames(xgb.pred_2018)[apply(xgb.pred_2018,1,which.max)]

preds_base_2018$max.prob <- max_prob_2018

preds_base_2018$pred.right <- ifelse(preds_base_2018$max.prob == "Out" & preds_base_2018$outcome_ec == 0, "yes", 
                                ifelse(preds_base_2018$max.prob == "Single" & preds_base_2018$outcome_ec == 1, "yes",
                                       ifelse(preds_base_2018$max.prob == "Double" & preds_base_2018$outcome_ec == 2, "yes",
                                              ifelse(preds_base_2018$max.prob == "Triple" & preds_base_2018$outcome_ec == 3, "yes",
                                                     ifelse(preds_base_2018$max.prob == "Home_Run" & preds_base_2018$outcome_ec == 4, "yes", "no")))))
preds_base_2018$pred.right <- as.factor(preds_base_2018$pred.right)

sum(preds_base_2018$pred.right == "yes") / nrow(preds_base_2018) #0.8160729

preds_base_2018 %>%
  group_by(outcome_ec, pred.right) %>%
  count()

###### 2019
xgb.pred_2019 = predict(base_mod_2019, x_test_2019, reshape=T)
xgb.pred_2019 = as.data.frame(xgb.pred_2019)
colnames(xgb.pred_2019) = c("Out", "Single", "Double", "Triple", "Home_Run")

preds_base_2019 <- test_2019
preds_base_2019$out.prob <- xgb.pred_2019[,1]
preds_base_2019$single.prob <- xgb.pred_2019[,2]
preds_base_2019$double.prob <- xgb.pred_2019[,3]
preds_base_2019$triple.prob <- xgb.pred_2019[,4]
preds_base_2019$hr.prob <- xgb.pred_2019[,5]

max_prob_2019 <- colnames(xgb.pred_2019)[apply(xgb.pred_2019,1,which.max)]

preds_base_2019$max.prob <- max_prob_2019

preds_base_2019$pred.right <- ifelse(preds_base_2019$max.prob == "Out" & preds_base_2019$outcome_ec == 0, "yes", 
                                ifelse(preds_base_2019$max.prob == "Single" & preds_base_2019$outcome_ec == 1, "yes",
                                       ifelse(preds_base_2019$max.prob == "Double" & preds_base_2019$outcome_ec == 2, "yes",
                                              ifelse(preds_base_2019$max.prob == "Triple" & preds_base_2019$outcome_ec == 3, "yes",
                                                     ifelse(preds_base_2019$max.prob == "Home_Run" & preds_base_2019$outcome_ec == 4, "yes", "no")))))
preds_base_2019$pred.right <- as.factor(preds_base_2019$pred.right)

sum(preds_base_2019$pred.right == "yes") / nrow(preds_base_2019) #0.8084982

preds_base_2019 %>%
  group_by(outcome_ec, pred.right) %>%
  count()
```


Now, I'm going to train a model on each entire season.
```{r}
### 2015
vec_label_2015 <- batted_ball_model_2015$outcome_ec
all_dataset_2015 <- as.matrix(batted_ball_model_2015 %>% dplyr::select(c("launch_angle", "launch_speed", "spray_angle", "stand_r", "p_throws_r", "temperature", "is_dome",
                                                            pitch_type_CH:venue_batter.stand_YankeeStadium_r)))
full_dataset_x_2015 = xgb.DMatrix(data = all_dataset_2015, label = vec_label_2015)

full_cv_2015 <- xgboost::xgb.cv(params = list(objective = "multi:softprob", eval_metric = c("mlogloss"), num_class = 5), data = full_dataset_x_2015, nrounds = 2500, early_stopping_rounds = 10, nfold=10, stratified = T)
full_mod_2015 <- xgboost::xgboost(params = list(objective = "multi:softprob", eval_metric = c("mlogloss"), num_class = 5), data = full_dataset_x_2015, nrounds = full_cv_2015$best_ntreelimit)
```

```{r include=FALSE}
### 2016
vec_label_2016 <- batted_ball_model_2016$outcome_ec
all_dataset_2016 <- as.matrix(batted_ball_model_2016 %>% dplyr::select(c("launch_angle", "launch_speed", "spray_angle", "stand_r", "p_throws_r", "temperature", "is_dome",
                                                            pitch_type_CH:venue_batter.stand_YankeeStadium_r)))
full_dataset_x_2016 = xgb.DMatrix(data = all_dataset_2016, label = vec_label_2016)

full_cv_2016 <- xgboost::xgb.cv(params = list(objective = "multi:softprob", eval_metric = c("mlogloss"), num_class = 5), data = full_dataset_x_2016, nrounds = 2500, early_stopping_rounds = 10, nfold=10, stratified = T)
full_mod_2016 <- xgboost::xgboost(params = list(objective = "multi:softprob", eval_metric = c("mlogloss"), num_class = 5), data = full_dataset_x_2016, nrounds = full_cv_2016$best_ntreelimit)

### 2017
vec_label_2017 <- batted_ball_model_2017$outcome_ec
all_dataset_2017 <- as.matrix(batted_ball_model_2017 %>% dplyr::select(c("launch_angle", "launch_speed", "spray_angle", "stand_r", "p_throws_r", "temperature", "is_dome",
                                                            pitch_type_CH:venue_batter.stand_YankeeStadium_r)))
full_dataset_x_2017 = xgb.DMatrix(data = all_dataset_2017, label = vec_label_2017)

full_cv_2017 <- xgboost::xgb.cv(params = list(objective = "multi:softprob", eval_metric = c("mlogloss"), num_class = 5), data = full_dataset_x_2017, nrounds = 2500, early_stopping_rounds = 10, nfold=10, stratified = T)
full_mod_2017 <- xgboost::xgboost(params = list(objective = "multi:softprob", eval_metric = c("mlogloss"), num_class = 5), data = full_dataset_x_2017, nrounds = full_cv_2017$best_ntreelimit)

### 2018
vec_label_2018 <- batted_ball_model_2018$outcome_ec
all_dataset_2018 <- as.matrix(batted_ball_model_2018 %>% dplyr::select(c("launch_angle", "launch_speed", "spray_angle", "stand_r", "p_throws_r", "temperature", "is_dome",
                                                            pitch_type_CH:venue_batter.stand_YankeeStadium_r)))
full_dataset_x_2018 = xgb.DMatrix(data = all_dataset_2018, label = vec_label_2018)

full_cv_2018 <- xgboost::xgb.cv(params = list(objective = "multi:softprob", eval_metric = c("mlogloss"), num_class = 5), data = full_dataset_x_2018, nrounds = 2500, early_stopping_rounds = 10, nfold=10, stratified = T)
full_mod_2018 <- xgboost::xgboost(params = list(objective = "multi:softprob", eval_metric = c("mlogloss"), num_class = 5), data = full_dataset_x_2018, nrounds = full_cv_2018$best_ntreelimit)

### 2019
vec_label_2019 <- batted_ball_model_2019$outcome_ec
all_dataset_2019 <- as.matrix(batted_ball_model_2019 %>% dplyr::select(c("launch_angle", "launch_speed", "spray_angle", "stand_r", "p_throws_r", "temperature", "is_dome",
                                                            pitch_type_CH:venue_batter.stand_YankeeStadium_r)))
full_dataset_x_2019 = xgb.DMatrix(data = all_dataset_2019, label = vec_label_2019)

full_cv_2019 <- xgboost::xgb.cv(params = list(objective = "multi:softprob", eval_metric = c("mlogloss"), num_class = 5), data = full_dataset_x_2019, nrounds = 2500, early_stopping_rounds = 10, nfold=10, stratified = T)
full_mod_2019 <- xgboost::xgboost(params = list(objective = "multi:softprob", eval_metric = c("mlogloss"), num_class = 5), data = full_dataset_x_2019, nrounds = full_cv_2019$best_ntreelimit)
```


Going to try and combine the dummy cols
```{r eval=FALSE, include=FALSE}
importance_full_mod_2019 <- xgb.importance(feature_names = colnames(full_mod_2019), model = full_mod_2019)

importance_plot_full_mod_2019 <- xgb.ggplot.importance(importance_matrix = importance_full_mod_2019)

grp_imp_base <- importance_full_mod_2019 %>%
  mutate(zone = case_when(grepl("zone", Feature) ~ "zone",
                          TRUE ~ "No")) %>%
  mutate(pitch_type = case_when(grepl("pitch_type", Feature) ~ "pitch_type",
                          TRUE ~ "No")) %>%
  mutate(batter_if_shift = case_when(grepl("batter_if_shift", Feature) ~ "batter_if_shift",
                          TRUE ~ "No")) %>%
  mutate(batter_of_shift = case_when(grepl("batter_of_shift", Feature) ~ "batter_of_shift",
                          TRUE ~ "No")) %>%
  mutate(game_year = case_when(grepl("game_year", Feature) ~ "game_year",
                          TRUE ~ "No")) %>%
  mutate(count = case_when(grepl("count", Feature) ~ "count",
                          TRUE ~ "No")) %>%
  mutate(venue_batter.stand = case_when(grepl("venue_batter.stand", Feature) ~ "venue_batter.stand",
                          TRUE ~ "No")) %>%
  mutate(batter_pitcher_matchup = case_when(grepl("batter_pitcher_matchup", Feature) ~ "batter_pitcher_matchup",
                          TRUE ~ "No"))

zone_imp <- grp_imp_base %>%
  group_by(zone) %>%
  filter(zone == "zone") %>%
  summarise(new_gain = sum(Gain), 
            new_cover = sum(Cover),
            new_frequency = sum(Frequency),
            new_importance = sum(Importance)) %>%
  rename(Feature = zone)

pitch_type_imp <- grp_imp_base %>%
  group_by(pitch_type) %>%
  filter(pitch_type == "pitch_type") %>%
  summarise(new_gain = sum(Gain), 
            new_cover = sum(Cover),
            new_frequency = sum(Frequency),
            new_importance = sum(Importance)) %>%
  rename(Feature = pitch_type)

batter_if_shift_imp <- grp_imp_base %>%
  group_by(batter_if_shift) %>%
  filter(batter_if_shift == "batter_if_shift") %>%
  summarise(new_gain = sum(Gain), 
            new_cover = sum(Cover),
            new_frequency = sum(Frequency),
            new_importance = sum(Importance)) %>%
  rename(Feature = batter_if_shift)

batter_of_shift_imp <- grp_imp_base %>%
  group_by(batter_of_shift) %>%
  filter(batter_of_shift == "batter_of_shift") %>%
  summarise(new_gain = sum(Gain), 
            new_cover = sum(Cover),
            new_frequency = sum(Frequency),
            new_importance = sum(Importance)) %>%
  rename(Feature = batter_of_shift)

game_year_imp <- grp_imp_base %>%
  group_by(game_year) %>%
  filter(game_year == "game_year") %>%
  summarise(new_gain = sum(Gain), 
            new_cover = sum(Cover),
            new_frequency = sum(Frequency),
            new_importance = sum(Importance)) %>%
  rename(Feature = game_year)

count_imp <- grp_imp_base %>%
  group_by(count) %>%
  filter(count == "count") %>%
  summarise(new_gain = sum(Gain), 
            new_cover = sum(Cover),
            new_frequency = sum(Frequency),
            new_importance = sum(Importance)) %>%
  rename(Feature = count)

venue_batter.stand_imp <- grp_imp_base %>%
  group_by(venue_batter.stand) %>%
  filter(venue_batter.stand == "venue_batter.stand") %>%
  summarise(new_gain = sum(Gain), 
            new_cover = sum(Cover),
            new_frequency = sum(Frequency),
            new_importance = sum(Importance)) %>%
  rename(Feature = venue_batter.stand)

batter_pitcher_matchup_imp <- grp_imp_base %>%
  group_by(batter_pitcher_matchup) %>%
  filter(batter_pitcher_matchup == "batter_pitcher_matchup") %>%
  summarise(new_gain = sum(Gain), 
            new_cover = sum(Cover),
            new_frequency = sum(Frequency),
            new_importance = sum(Importance)) %>%
  rename(Feature = batter_pitcher_matchup)

new_imp <- rbind(zone_imp, pitch_type_imp, batter_if_shift_imp, batter_of_shift_imp, game_year_imp, count_imp, venue_batter.stand_imp, 
                 batter_pitcher_matchup_imp)

new_imp <- rename(new_imp, Gain = new_gain, Cover = new_cover, Frequency = new_frequency, Importance = new_importance)

non_dummies <- importance_base_mod_2019 %>%
  filter(Feature == "launch_angle" | Feature == "launch_speed" | Feature == "spray_angle" | Feature == "stand_r" | Feature == "temperature" | 
           Feature == "p_throws_r" | Feature == "is_dome")

all_grp_importance <- rbind(non_dummies, new_imp)

grp_importance_plot_full_mod_2019 <- xgb.ggplot.importance(importance_matrix = all_grp_importance)
```

Feature importance plots.
```{r}
final_2015 <- grp_importance_plot_full_mod_2015 + ggtitle("Feature Importance - 2015 Model")

final_2016 <- grp_importance_plot_full_mod_2016 + ggtitle("Feature Importance - 2016 Model")

final_2017 <- grp_importance_plot_full_mod_2017 + ggtitle("Feature Importance - 2017 Model")

final_2018 <- grp_importance_plot_full_mod_2018 + ggtitle("Feature Importance - 2018 Model")

final_2019 <- grp_importance_plot_full_mod_2019 + ggtitle("Feature Importance - 2019 Model")

(final_2015 + final_2016 + final_2017) / (final_2018 + final_2019)
```

Full model predictions. These are pretty great; ranging from 85.25% to 86.25%. 
```{r}
xgb.full_pred_2015 = predict(full_mod_2015, full_dataset_x_2015, reshape=T)
xgb.full_pred_2015 = as.data.frame(xgb.full_pred_2015)
colnames(xgb.full_pred_2015) = c("Out", "Single", "Double", "Triple", "Home_Run")

preds_full_2015 <- batted_ball_model_2015
preds_full_2015$out.prob <- xgb.full_pred_2015[,1]
preds_full_2015$single.prob <- xgb.full_pred_2015[,2]
preds_full_2015$double.prob <- xgb.full_pred_2015[,3]
preds_full_2015$triple.prob <- xgb.full_pred_2015[,4]
preds_full_2015$hr.prob <- xgb.full_pred_2015[,5]

max_prob_full_2015 <- colnames(xgb.full_pred_2015)[apply(xgb.full_pred_2015,1,which.max)]

preds_full_2015$max.prob <- max_prob_full_2015

preds_full_2015$pred.right <- ifelse(preds_full_2015$max.prob == "Out" & preds_full_2015$outcome_ec == 0, "yes", 
                                ifelse(preds_full_2015$max.prob == "Single" & preds_full_2015$outcome_ec == 1, "yes",
                                       ifelse(preds_full_2015$max.prob == "Double" & preds_full_2015$outcome_ec == 2, "yes",
                                              ifelse(preds_full_2015$max.prob == "Triple" & preds_full_2015$outcome_ec == 3, "yes",
                                                     ifelse(preds_full_2015$max.prob == "Home_Run" & preds_full_2015$outcome_ec == 4, "yes", "no")))))
preds_full_2015$pred.right <- as.factor(preds_full_2015$pred.right)

sum(preds_full_2015$pred.right == "yes") / nrow(preds_full_2015) #0.8544968

preds_full_2015 %>%
  group_by(outcome_ec, pred.right) %>%
  count()

###### 2016
xgb.full_pred_2016 = predict(full_mod_2016, full_dataset_x_2016, reshape=T)
xgb.full_pred_2016 = as.data.frame(xgb.full_pred_2016)
colnames(xgb.full_pred_2016) = c("Out", "Single", "Double", "Triple", "Home_Run")

preds_full_2016 <- batted_ball_model_2016
preds_full_2016$out.prob <- xgb.full_pred_2016[,1]
preds_full_2016$single.prob <- xgb.full_pred_2016[,2]
preds_full_2016$double.prob <- xgb.full_pred_2016[,3]
preds_full_2016$triple.prob <- xgb.full_pred_2016[,4]
preds_full_2016$hr.prob <- xgb.full_pred_2016[,5]

max_prob_full_2016 <- colnames(xgb.full_pred_2016)[apply(xgb.full_pred_2016,1,which.max)]

preds_full_2016$max.prob <- max_prob_full_2016

preds_full_2016$pred.right <- ifelse(preds_full_2016$max.prob == "Out" & preds_full_2016$outcome_ec == 0, "yes", 
                                ifelse(preds_full_2016$max.prob == "Single" & preds_full_2016$outcome_ec == 1, "yes",
                                       ifelse(preds_full_2016$max.prob == "Double" & preds_full_2016$outcome_ec == 2, "yes",
                                              ifelse(preds_full_2016$max.prob == "Triple" & preds_full_2016$outcome_ec == 3, "yes",
                                                     ifelse(preds_full_2016$max.prob == "Home_Run" & preds_full_2016$outcome_ec == 4, "yes", "no")))))
preds_full_2016$pred.right <- as.factor(preds_full_2016$pred.right)

sum(preds_full_2016$pred.right == "yes") / nrow(preds_full_2016) #0.852469

preds_full_2016 %>%
  group_by(outcome_ec, pred.right) %>%
  count()

###### 2017
xgb.full_pred_2017 = predict(full_mod_2017, full_dataset_x_2017, reshape=T)
xgb.full_pred_2017 = as.data.frame(xgb.full_pred_2017)
colnames(xgb.full_pred_2017) = c("Out", "Single", "Double", "Triple", "Home_Run")

preds_full_2017 <- batted_ball_model_2017
preds_full_2017$out.prob <- xgb.full_pred_2017[,1]
preds_full_2017$single.prob <- xgb.full_pred_2017[,2]
preds_full_2017$double.prob <- xgb.full_pred_2017[,3]
preds_full_2017$triple.prob <- xgb.full_pred_2017[,4]
preds_full_2017$hr.prob <- xgb.full_pred_2017[,5]

max_prob_full_2017 <- colnames(xgb.full_pred_2017)[apply(xgb.full_pred_2017,1,which.max)]

preds_full_2017$max.prob <- max_prob_full_2017

preds_full_2017$pred.right <- ifelse(preds_full_2017$max.prob == "Out" & preds_full_2017$outcome_ec == 0, "yes", 
                                ifelse(preds_full_2017$max.prob == "Single" & preds_full_2017$outcome_ec == 1, "yes",
                                       ifelse(preds_full_2017$max.prob == "Double" & preds_full_2017$outcome_ec == 2, "yes",
                                              ifelse(preds_full_2017$max.prob == "Triple" & preds_full_2017$outcome_ec == 3, "yes",
                                                     ifelse(preds_full_2017$max.prob == "Home_Run" & preds_full_2017$outcome_ec == 4, "yes", "no")))))
preds_full_2017$pred.right <- as.factor(preds_full_2017$pred.right)

sum(preds_full_2017$pred.right == "yes") / nrow(preds_full_2017) #0.8606125

preds_full_2017 %>%
  group_by(outcome_ec, pred.right) %>%
  count()

###### 2018
xgb.full_pred_2018 = predict(full_mod_2018, full_dataset_x_2018, reshape=T)
xgb.full_pred_2018 = as.data.frame(xgb.full_pred_2018)
colnames(xgb.full_pred_2018) = c("Out", "Single", "Double", "Triple", "Home_Run")

preds_full_2018 <- batted_ball_model_2018
preds_full_2018$out.prob <- xgb.full_pred_2018[,1]
preds_full_2018$single.prob <- xgb.full_pred_2018[,2]
preds_full_2018$double.prob <- xgb.full_pred_2018[,3]
preds_full_2018$triple.prob <- xgb.full_pred_2018[,4]
preds_full_2018$hr.prob <- xgb.full_pred_2018[,5]

max_prob_full_2018 <- colnames(xgb.full_pred_2018)[apply(xgb.full_pred_2018,1,which.max)]

preds_full_2018$max.prob <- max_prob_full_2018

preds_full_2018$pred.right <- ifelse(preds_full_2018$max.prob == "Out" & preds_full_2018$outcome_ec == 0, "yes", 
                                ifelse(preds_full_2018$max.prob == "Single" & preds_full_2018$outcome_ec == 1, "yes",
                                       ifelse(preds_full_2018$max.prob == "Double" & preds_full_2018$outcome_ec == 2, "yes",
                                              ifelse(preds_full_2018$max.prob == "Triple" & preds_full_2018$outcome_ec == 3, "yes",
                                                     ifelse(preds_full_2018$max.prob == "Home_Run" & preds_full_2018$outcome_ec == 4, "yes", "no")))))
preds_full_2018$pred.right <- as.factor(preds_full_2018$pred.right)

sum(preds_full_2018$pred.right == "yes") / nrow(preds_full_2018) #0.8562713

preds_full_2018 %>%
  group_by(outcome_ec, pred.right) %>%
  count()

###### 2019
xgb.full_pred_2019 = predict(full_mod_2019, full_dataset_x_2019, reshape=T)
xgb.full_pred_2019 = as.data.frame(xgb.full_pred_2019)
colnames(xgb.full_pred_2019) = c("Out", "Single", "Double", "Triple", "Home_Run")

preds_full_2019 <- batted_ball_model_2019
preds_full_2019$out.prob <- xgb.full_pred_2019[,1]
preds_full_2019$single.prob <- xgb.full_pred_2019[,2]
preds_full_2019$double.prob <- xgb.full_pred_2019[,3]
preds_full_2019$triple.prob <- xgb.full_pred_2019[,4]
preds_full_2019$hr.prob <- xgb.full_pred_2019[,5]

max_prob_full_2019 <- colnames(xgb.full_pred_2019)[apply(xgb.full_pred_2019,1,which.max)]

preds_full_2019$max.prob <- max_prob_full_2019

preds_full_2019$pred.right <- ifelse(preds_full_2019$max.prob == "Out" & preds_full_2019$outcome_ec == 0, "yes", 
                                ifelse(preds_full_2019$max.prob == "Single" & preds_full_2019$outcome_ec == 1, "yes",
                                       ifelse(preds_full_2019$max.prob == "Double" & preds_full_2019$outcome_ec == 2, "yes",
                                              ifelse(preds_full_2019$max.prob == "Triple" & preds_full_2019$outcome_ec == 3, "yes",
                                                     ifelse(preds_full_2019$max.prob == "Home_Run" & preds_full_2019$outcome_ec == 4, "yes", "no")))))
preds_full_2019$pred.right <- as.factor(preds_full_2019$pred.right)

sum(preds_full_2019$pred.right == "yes") / nrow(preds_full_2019) #0.862459

preds_full_2019 %>%
  group_by(outcome_ec, pred.right) %>%
  count()
```

```{r eval=FALSE, include=FALSE}
library(janitor)
preds_full_2015 %>%
  group_by(outcome_ec, pred.right) %>%
  mutate(outcome_w_pred.right = paste0(outcome_ec, "_", pred.right)) %>%
  ungroup() %>%
  filter(outcome_ec == 3) %>%
  tabyl(outcome_w_pred.right) %>%
  adorn_totals("row")

#     2015 xgb model                   
#     0_no  4715 0.06859579
#     0_yes 64021 0.93140421
#     1_no  6636 0.2885719
#     1_yes 16360 0.7114281
#     2_no 2780 0.4062546
#     2_yes 4063 0.5937454
#     3_no 618 0.7923077
#     3_yes 162 0.2076923
#     4_no  357 0.07997312
#     4_yes 4107 0.92002688

#   2015 glmer model
#   0_no  6224 0.09054935
#   0_yes 62512 0.90945065
#   1_no  6036 0.2624804
#   1_yes 16960 0.7375196
#   2_no 2121 0.3099518
#   2_yes 4722 0.6900482
#   3_no 300 0.3846154
#   3_yes 480 0.6153846
#   4_no  306 0.06854839
#   4_yes 4158 0.93145161

model.coef.2015 %>%
  group_by(outcome_ec, pred.right_new) %>%
  mutate(outcome_w_pred.right = paste0(outcome_ec, "_", pred.right_new)) %>%
  ungroup() %>%
  filter(outcome_ec == 4) %>%
  tabyl(outcome_w_pred.right) %>%
  adorn_totals("row")

#   2015 glmer model
#   0_no  6224 0.09054935
#   0_yes 62512 0.90945065
#   1_no  6036 0.2624804
#   1_yes 16960 0.7375196
#   2_no 2121 0.3099518
#   2_yes 4722 0.6900482
#   3_no 300 0.3846154
#   3_yes 480 0.6153846
#   4_no  306 0.06854839
#   4_yes 4158 0.93145161

```


Now, the next step is to create the generalized linear mixed-effects model building off of the XGBoost models for each season.

```{r}
df.single <- preds_full_2015 %>%
  filter(outcome_ec == 0 | outcome_ec == 1) %>%
  mutate(outcome = ifelse(outcome_ec == 0, 0, 1))
df.double <- preds_full_2015 %>%
  filter(outcome_ec == 0 | outcome_ec == 2) %>%
  mutate(outcome = ifelse(outcome_ec == 0, 0, 1))
df.triple <- preds_full_2015 %>%
  filter(outcome_ec == 0 | outcome_ec == 3) %>%
  mutate(outcome = ifelse(outcome_ec == 0, 0, 1))
df.home_run <- preds_full_2015 %>%
  filter(outcome_ec == 0 | outcome_ec == 4) %>%
  mutate(outcome = ifelse(outcome_ec == 0, 0, 1))

glmer.single.mod.2015 <- glmer(
  outcome ~
    (1|pitcher_name) +
    (1|player_name) +
    logit(single.prob),
  data=df.single,
  family=binomial(link = "probit"),
  nAGQ=0)
glmer.double.mod.2015 <- glmer(
  outcome ~
    (1|pitcher_name) +
    (1|player_name) +     
    logit(double.prob),
  data=df.double,
  family=binomial(link = "probit"),
  nAGQ=0)
glmer.triple.mod.2015 <- glmer(
  outcome ~
    (1|pitcher_name) +
    (1|player_name) +
    logit(triple.prob),
  data=df.triple,
  family=binomial(link = "probit"),
  nAGQ=0)
glmer.home_run.mod.2015 <- glmer(
  outcome ~
    (1|pitcher_name) +
    (1|player_name) + 
    logit(hr.prob),
  data=df.home_run,
  family=binomial(link = "probit"),
  nAGQ=0)

df.preds_2015 <- preds_full_2015
df.preds_2015$single_all_lp_bat <- predict(glmer.single.mod.2015, newdata=df.preds_2015,
                                    allow.new.levels = TRUE, type='link', re.form=~(1|player_name))
df.preds_2015$double_all_lp_bat <- predict(glmer.double.mod.2015, newdata=df.preds_2015,
                                    allow.new.levels = TRUE, type='link', re.form=~(1|player_name))
df.preds_2015$triple_all_lp_bat <- predict(glmer.triple.mod.2015, newdata=df.preds_2015,
                                    allow.new.levels = TRUE, type='link', re.form=~(1|player_name))
df.preds_2015$home_run_all_lp_bat <- predict(glmer.home_run.mod.2015, newdata=df.preds_2015,
                                    allow.new.levels = TRUE, type='link', re.form=~(1|player_name))

model.coef.2015 <- df.preds_2015 %>%
  dplyr::mutate(
    K_all = 1 + exp(single_all_lp_bat) + exp(double_all_lp_bat) + exp(triple_all_lp_bat) + exp(home_run_all_lp_bat),
    outs_new.prob = 1 / K_all,
    single_new.prob = exp(single_all_lp_bat) / K_all,
    double_new.prob = exp(double_all_lp_bat) / K_all,
    triple_new.prob = exp(triple_all_lp_bat) / K_all,
    home_run_new.prob = exp(home_run_all_lp_bat) / K_all)

model.coef.2015 %<>%
  mutate(max.prob_ec = ifelse(outs_new.prob > single_new.prob & outs_new.prob > double_new.prob & outs_new.prob > triple_new.prob & outs_new.prob > home_run_new.prob, 0,
                           ifelse(single_new.prob > outs_new.prob & single_new.prob > double_new.prob & single_new.prob > triple_new.prob & 
                                    single_new.prob > home_run_new.prob, 1,
                                  ifelse(double_new.prob > outs_new.prob & double_new.prob > single_new.prob & double_new.prob > triple_new.prob 
                                         & double_new.prob > home_run_new.prob, 2,
                                         ifelse(triple_new.prob > outs_new.prob & triple_new.prob > single_new.prob & triple_new.prob > double_new.prob &
                                                  triple_new.prob > home_run_new.prob,  3, 4)))))

model.coef.2015 %<>%
  mutate(pred.right_new = ifelse(max.prob_ec == outcome_ec, "yes", "no"))

sum(model.coef.2015$pred.right_new == "yes") / nrow((model.coef.2015)) #  xgb is 0.8544968 this is 0.855643
```


```{r include=FALSE}
############ 2016
df.single <- preds_full_2016 %>%
  filter(outcome_ec == 0 | outcome_ec == 1) %>%
  mutate(outcome = ifelse(outcome_ec == 0, 0, 1))
df.double <- preds_full_2016 %>%
  filter(outcome_ec == 0 | outcome_ec == 2) %>%
  mutate(outcome = ifelse(outcome_ec == 0, 0, 1))
df.triple <- preds_full_2016 %>%
  filter(outcome_ec == 0 | outcome_ec == 3) %>%
  mutate(outcome = ifelse(outcome_ec == 0, 0, 1))
df.home_run <- preds_full_2016 %>%
  filter(outcome_ec == 0 | outcome_ec == 4) %>%
  mutate(outcome = ifelse(outcome_ec == 0, 0, 1))

glmer.single.mod.2016 <- glmer(
  outcome ~
    (1|pitcher_name) + (1|player_name) + logit(single.prob),
  data=df.single,
  family=binomial(link = "probit"),
  nAGQ=0)
glmer.double.mod.2016 <- glmer(
  outcome ~
    (1|pitcher_name) + (1|player_name) + logit(double.prob),
  data=df.double,
  family=binomial(link = "probit"),
  nAGQ=0)
glmer.triple.mod.2016 <- glmer(
  outcome ~
    (1|pitcher_name) + (1|player_name) + logit(triple.prob),
  data=df.triple,
  family=binomial(link = "probit"),
  nAGQ=0)
glmer.home_run.mod.2016 <- glmer(
  outcome ~
    (1|pitcher_name) + (1|player_name) + logit(hr.prob),
  data=df.home_run,
  family=binomial(link = "probit"),
  nAGQ=0)

df.preds_2016 <- preds_full_2016
df.preds_2016$single_all_lp_bat <- predict(glmer.single.mod.2016, newdata=df.preds_2016,
                                    allow.new.levels = TRUE, type='link', re.form=~(1|player_name))
df.preds_2016$double_all_lp_bat <- predict(glmer.double.mod.2016, newdata=df.preds_2016,
                                    allow.new.levels = TRUE, type='link', re.form=~(1|player_name))
df.preds_2016$triple_all_lp_bat <- predict(glmer.triple.mod.2016, newdata=df.preds_2016,
                                    allow.new.levels = TRUE, type='link', re.form=~(1|player_name))
df.preds_2016$home_run_all_lp_bat <- predict(glmer.home_run.mod.2016, newdata=df.preds_2016,
                                    allow.new.levels = TRUE, type='link', re.form=~(1|player_name))
model.coef.2016 <- df.preds_2016 %>%
  dplyr::mutate(
    K_all = 1 + exp(single_all_lp_bat) + exp(double_all_lp_bat) + exp(triple_all_lp_bat) + exp(home_run_all_lp_bat),
    outs_new.prob = 1 / K_all,
    single_new.prob = exp(single_all_lp_bat) / K_all,
    double_new.prob = exp(double_all_lp_bat) / K_all,
    triple_new.prob = exp(triple_all_lp_bat) / K_all,
    home_run_new.prob = exp(home_run_all_lp_bat) / K_all
  )

model.coef.2016 %<>%
  mutate(max.prob_ec = ifelse(outs_new.prob > single_new.prob & outs_new.prob > double_new.prob & outs_new.prob > triple_new.prob & 
                                outs_new.prob > home_run_new.prob, 0,
                           ifelse(single_new.prob > outs_new.prob & single_new.prob > double_new.prob & single_new.prob > triple_new.prob & 
                                    single_new.prob > home_run_new.prob, 1,
                                  ifelse(double_new.prob > outs_new.prob & double_new.prob > single_new.prob & double_new.prob > triple_new.prob 
                                         & double_new.prob > home_run_new.prob, 2,
                                         ifelse(triple_new.prob > outs_new.prob & triple_new.prob > single_new.prob & triple_new.prob > double_new.prob &
                                                  triple_new.prob > home_run_new.prob, 3, 4)))))

model.coef.2016 %<>%
  mutate(pred.right_new = ifelse(max.prob_ec == outcome_ec, "yes", "no"))

sum(model.coef.2016$pred.right_new == "yes") / nrow((model.coef.2016)) #  xgb is 0.852469 this is 0.8528027


############ 2017
df.single <- preds_full_2017 %>%
  filter(outcome_ec == 0 | outcome_ec == 1) %>%
  mutate(outcome = ifelse(outcome_ec == 0, 0, 1))
df.double <- preds_full_2017 %>%
  filter(outcome_ec == 0 | outcome_ec == 2) %>%
  mutate(outcome = ifelse(outcome_ec == 0, 0, 1))
df.triple <- preds_full_2017 %>%
  filter(outcome_ec == 0 | outcome_ec == 3) %>%
  mutate(outcome = ifelse(outcome_ec == 0, 0, 1))
df.home_run <- preds_full_2017 %>%
  filter(outcome_ec == 0 | outcome_ec == 4) %>%
  mutate(outcome = ifelse(outcome_ec == 0, 0, 1))

glmer.single.mod.2017 <- glmer(
  outcome ~
    (1|pitcher_name) + (1|player_name) + logit(single.prob),
  data=df.single,
  family=binomial(link = "probit"),
  nAGQ=0)
glmer.double.mod.2017 <- glmer(
  outcome ~
    (1|pitcher_name) + (1|player_name) + logit(double.prob),
  data=df.double,
  family=binomial(link = "probit"),
  nAGQ=0)
glmer.triple.mod.2017 <- glmer(
  outcome ~
    (1|pitcher_name) + (1|player_name) + logit(triple.prob),
  data=df.triple,
  family=binomial(link = "probit"),
  nAGQ=0)
glmer.home_run.mod.2017 <- glmer(
  outcome ~
    (1|pitcher_name) + (1|player_name) + logit(hr.prob),
  data=df.home_run,
  family=binomial(link = "probit"),
  nAGQ=0)

df.preds_2017 <- preds_full_2017
df.preds_2017$single_all_lp_bat <- predict(glmer.single.mod.2017, newdata=df.preds_2017,
                                    allow.new.levels = TRUE, type='link', re.form=~(1|player_name))
df.preds_2017$double_all_lp_bat <- predict(glmer.double.mod.2017, newdata=df.preds_2017,
                                    allow.new.levels = TRUE, type='link', re.form=~(1|player_name))
df.preds_2017$triple_all_lp_bat <- predict(glmer.triple.mod.2017, newdata=df.preds_2017,
                                    allow.new.levels = TRUE, type='link', re.form=~(1|player_name))
df.preds_2017$home_run_all_lp_bat <- predict(glmer.home_run.mod.2017, newdata=df.preds_2017,
                                    allow.new.levels = TRUE, type='link', re.form=~(1|player_name))
model.coef.2017 <- df.preds_2017 %>%
  dplyr::mutate(
    K_all = 1 + exp(single_all_lp_bat) + exp(double_all_lp_bat) + exp(triple_all_lp_bat) + exp(home_run_all_lp_bat),
    outs_new.prob = 1 / K_all,
    single_new.prob = exp(single_all_lp_bat) / K_all,
    double_new.prob = exp(double_all_lp_bat) / K_all,
    triple_new.prob = exp(triple_all_lp_bat) / K_all,
    home_run_new.prob = exp(home_run_all_lp_bat) / K_all
  )

model.coef.2017 %<>%
  mutate(max.prob_ec = ifelse(outs_new.prob > single_new.prob & outs_new.prob > double_new.prob & outs_new.prob > triple_new.prob & 
                                outs_new.prob > home_run_new.prob, 0,
                           ifelse(single_new.prob > outs_new.prob & single_new.prob > double_new.prob & single_new.prob > triple_new.prob & 
                                    single_new.prob > home_run_new.prob, 1,
                                  ifelse(double_new.prob > outs_new.prob & double_new.prob > single_new.prob & double_new.prob > triple_new.prob 
                                         & double_new.prob > home_run_new.prob, 2,
                                         ifelse(triple_new.prob > outs_new.prob & triple_new.prob > single_new.prob & triple_new.prob > double_new.prob &
                                                  triple_new.prob > home_run_new.prob, 3, 4)))))

model.coef.2017 %<>%
  mutate(pred.right_new = ifelse(max.prob_ec == outcome_ec, "yes", "no"))

sum(model.coef.2017$pred.right_new == "yes") / nrow((model.coef.2017)) # xgb is 0.8606125 this is 0.8621587


############ 2018
df.single <- preds_full_2018 %>%
  filter(outcome_ec == 0 | outcome_ec == 1) %>%
  mutate(outcome = ifelse(outcome_ec == 0, 0, 1))
df.double <- preds_full_2018 %>%
  filter(outcome_ec == 0 | outcome_ec == 2) %>%
  mutate(outcome = ifelse(outcome_ec == 0, 0, 1))
df.triple <- preds_full_2018 %>%
  filter(outcome_ec == 0 | outcome_ec == 3) %>%
  mutate(outcome = ifelse(outcome_ec == 0, 0, 1))
df.home_run <- preds_full_2018 %>%
  filter(outcome_ec == 0 | outcome_ec == 4) %>%
  mutate(outcome = ifelse(outcome_ec == 0, 0, 1))

glmer.single.mod.2018 <- glmer(
  outcome ~
    (1|pitcher_name) + (1|player_name) + logit(single.prob),
  data=df.single,
  family=binomial(link = "probit"),
  nAGQ=0)
glmer.double.mod.2018 <- glmer(
  outcome ~
    (1|pitcher_name) + (1|player_name) + logit(double.prob),
  data=df.double,
  family=binomial(link = "probit"),
  nAGQ=0)
glmer.triple.mod.2018 <- glmer(
  outcome ~
    (1|pitcher_name) + (1|player_name) + logit(triple.prob),
  data=df.triple,
  family=binomial(link = "probit"),
  nAGQ=0)
glmer.home_run.mod.2018 <- glmer(
  outcome ~
    (1|pitcher_name) + (1|player_name) + logit(hr.prob),
  data=df.home_run,
  family=binomial(link = "probit"),
  nAGQ=0)

df.preds_2018 <- preds_full_2018
df.preds_2018$single_all_lp_bat <- predict(glmer.single.mod.2018, newdata=df.preds_2018,
                                    allow.new.levels = TRUE, type='link', re.form=~(1|player_name))
df.preds_2018$double_all_lp_bat <- predict(glmer.double.mod.2018, newdata=df.preds_2018,
                                    allow.new.levels = TRUE, type='link', re.form=~(1|player_name))
df.preds_2018$triple_all_lp_bat <- predict(glmer.triple.mod.2018, newdata=df.preds_2018,
                                    allow.new.levels = TRUE, type='link', re.form=~(1|player_name))
df.preds_2018$home_run_all_lp_bat <- predict(glmer.home_run.mod.2018, newdata=df.preds_2018,
                                    allow.new.levels = TRUE, type='link', re.form=~(1|player_name))
model.coef.2018 <- df.preds_2018 %>%
  dplyr::mutate(
    K_all = 1 + exp(single_all_lp_bat) + exp(double_all_lp_bat) + exp(triple_all_lp_bat) + exp(home_run_all_lp_bat),
    outs_new.prob = 1 / K_all,
    single_new.prob = exp(single_all_lp_bat) / K_all,
    double_new.prob = exp(double_all_lp_bat) / K_all,
    triple_new.prob = exp(triple_all_lp_bat) / K_all,
    home_run_new.prob = exp(home_run_all_lp_bat) / K_all
  )

model.coef.2018 %<>%
  mutate(max.prob_ec = ifelse(outs_new.prob > single_new.prob & outs_new.prob > double_new.prob & outs_new.prob > triple_new.prob & 
                                outs_new.prob > home_run_new.prob, 0,
                           ifelse(single_new.prob > outs_new.prob & single_new.prob > double_new.prob & single_new.prob > triple_new.prob & 
                                    single_new.prob > home_run_new.prob, 1,
                                  ifelse(double_new.prob > outs_new.prob & double_new.prob > single_new.prob & double_new.prob > triple_new.prob 
                                         & double_new.prob > home_run_new.prob, 2,
                                         ifelse(triple_new.prob > outs_new.prob & triple_new.prob > single_new.prob & triple_new.prob > double_new.prob &
                                                  triple_new.prob > home_run_new.prob, 3, 4)))))

model.coef.2018 %<>%
  mutate(pred.right_new = ifelse(max.prob_ec == outcome_ec, "yes", "no"))

sum(model.coef.2018$pred.right_new == "yes") / nrow((model.coef.2018)) # xgb is 0.8562713 this is 0.857305


############ 2019
df.single <- preds_full_2019 %>%
  filter(outcome_ec == 0 | outcome_ec == 1) %>%
  mutate(outcome = ifelse(outcome_ec == 0, 0, 1))
df.double <- preds_full_2019 %>%
  filter(outcome_ec == 0 | outcome_ec == 2) %>%
  mutate(outcome = ifelse(outcome_ec == 0, 0, 1))
df.triple <- preds_full_2019 %>%
  filter(outcome_ec == 0 | outcome_ec == 3) %>%
  mutate(outcome = ifelse(outcome_ec == 0, 0, 1))
df.home_run <- preds_full_2019 %>%
  filter(outcome_ec == 0 | outcome_ec == 4) %>%
  mutate(outcome = ifelse(outcome_ec == 0, 0, 1))

glmer.single.mod.2019 <- glmer(
  outcome ~
    (1|pitcher_name) + (1|player_name) + logit(single.prob),
  data=df.single,
  family=binomial(link = "probit"),
  nAGQ=0)
glmer.double.mod.2019 <- glmer(
  outcome ~
    (1|pitcher_name) + (1|player_name) + logit(double.prob),
  data=df.double,
  family=binomial(link = "probit"),
  nAGQ=0)
glmer.triple.mod.2019 <- glmer(
  outcome ~
    (1|pitcher_name) + (1|player_name) + logit(triple.prob),
  data=df.triple,
  family=binomial(link = "probit"),
  nAGQ=0)
glmer.home_run.mod.2019 <- glmer(
  outcome ~
    (1|pitcher_name) + (1|player_name) + logit(hr.prob),
  data=df.home_run,
  family=binomial(link = "probit"),
  nAGQ=0)

df.preds_2019 <- preds_full_2019
df.preds_2019$single_all_lp_bat <- predict(glmer.single.mod.2019, newdata=df.preds_2019,
                                    allow.new.levels = TRUE, type='link', re.form=~(1|player_name))
df.preds_2019$double_all_lp_bat <- predict(glmer.double.mod.2019, newdata=df.preds_2019,
                                    allow.new.levels = TRUE, type='link', re.form=~(1|player_name))
df.preds_2019$triple_all_lp_bat <- predict(glmer.triple.mod.2019, newdata=df.preds_2019,
                                    allow.new.levels = TRUE, type='link', re.form=~(1|player_name))
df.preds_2019$home_run_all_lp_bat <- predict(glmer.home_run.mod.2019, newdata=df.preds_2019,
                                    allow.new.levels = TRUE, type='link', re.form=~(1|player_name))
model.coef.2019 <- df.preds_2019 %>%
  dplyr::mutate(
    K_all = 1 + exp(single_all_lp_bat) + exp(double_all_lp_bat) + exp(triple_all_lp_bat) + exp(home_run_all_lp_bat),
    outs_new.prob = 1 / K_all,
    single_new.prob = exp(single_all_lp_bat) / K_all,
    double_new.prob = exp(double_all_lp_bat) / K_all,
    triple_new.prob = exp(triple_all_lp_bat) / K_all,
    home_run_new.prob = exp(home_run_all_lp_bat) / K_all
  )

model.coef.2019 %<>%
  mutate(max.prob_ec = ifelse(outs_new.prob > single_new.prob & outs_new.prob > double_new.prob & outs_new.prob > triple_new.prob & 
                                outs_new.prob > home_run_new.prob, 0,
                           ifelse(single_new.prob > outs_new.prob & single_new.prob > double_new.prob & single_new.prob > triple_new.prob & 
                                    single_new.prob > home_run_new.prob, 1,
                                  ifelse(double_new.prob > outs_new.prob & double_new.prob > single_new.prob & double_new.prob > triple_new.prob 
                                         & double_new.prob > home_run_new.prob, 2,
                                         ifelse(triple_new.prob > outs_new.prob & triple_new.prob > single_new.prob & triple_new.prob > double_new.prob &
                                                  triple_new.prob > home_run_new.prob, 3, 4)))))

model.coef.2019 %<>%
  mutate(pred.right_new = ifelse(max.prob_ec == outcome_ec, "yes", "no"))

sum(model.coef.2019$pred.right_new == "yes") / nrow((model.coef.2019)) # xgb is 0.862459 this is 0.8640512
```

```{r}
model.coef.2016 %>%
  mutate(new_old_diff_out = outs_new.prob - out.prob, 
         new_old_diff_single = single_new.prob - single.prob,
         new_old_diff_double = double_new.prob - double.prob,
         new_old_diff_triple = triple_new.prob - triple.prob,
         new_old_diff_hr = home_run_new.prob - hr.prob) %>%
  dplyr::select(c("player_name", "pitcher_name",   "out.prob", "outs_new.prob", "new_old_diff_out",  "single.prob", "new_old_diff_single",  "single_new.prob",
                  "double.prob", "double_new.prob", "new_old_diff_double",  "triple.prob", "triple_new.prob", "new_old_diff_triple", "hr.prob", "home_run_new.prob",
                  "new_old_diff_hr", "max.prob", "pred.right","game_date", "venue_batter.stand", "pitch_type", "outcome_ec", "launch_speed", "launch_angle", 
                  "spray_angle", "batter_if_shift", "batter_of_shift")) %>%
  View()

model.coef.2016 %>%
  mutate(new_old_diff_out = outs_new.prob - out.prob, 
         new_old_diff_single = single_new.prob - single.prob,
         new_old_diff_double = double_new.prob - double.prob,
         new_old_diff_triple = triple_new.prob - triple.prob,
         new_old_diff_hr = home_run_new.prob - hr.prob) %>% 
  summarise(mean_out_diff = var(new_old_diff_out),
            mean_single_diff = var(new_old_diff_single),
            mean_double_diff = var(new_old_diff_double),
            mean_triple_diff = var(new_old_diff_triple),
            mean_old_diff_hr = var(new_old_diff_hr))

model.coef.2019 %>%
  mutate(new_old_diff_out = outs_new.prob - out.prob, 
         new_old_diff_single = single_new.prob - single.prob,
         new_old_diff_double = double_new.prob - double.prob,
         new_old_diff_triple = triple_new.prob - triple.prob,
         new_old_diff_hr = home_run_new.prob - hr.prob) %>%
  dplyr::select(c("player_name", "pitcher_name",   "out.prob", "outs_new.prob", "new_old_diff_out",  "single.prob", "new_old_diff_single",  "single_new.prob",
                  "double.prob", "double_new.prob", "new_old_diff_double",  "triple.prob", "triple_new.prob", "new_old_diff_triple", "hr.prob", "home_run_new.prob",
                  "new_old_diff_hr", "max.prob", "pred.right","game_date", "venue_batter.stand", "pitch_type", "outcome_ec", "launch_speed", "launch_angle", 
                  "spray_angle", "batter_if_shift", "batter_of_shift")) %>%
  View()
```


Making QC+ metric
```{r}
woba_weights_2015 <- c(0, .881,	1.256, 1.594, 2.065) 

model.coef.2015 %<>%
  mutate(expected_woba = (outs_new.prob * woba_weights_2015[1]) + (single_new.prob * woba_weights_2015[2]) + (double_new.prob * woba_weights_2015[3]) + 
           (triple_new.prob * woba_weights_2015[4]) + (home_run_new.prob * woba_weights_2015[5]))

model.coef.2015$percentile <- 100 * (scales::rescale(model.coef.2015$expected_woba, to=c(0,1)))
model.coef.2015$qcp <- psych::rescale(model.coef.2015$percentile, mean = 100, sd = 30, df=F)
```

```{r include=FALSE}
woba_weights_2016 <- c(0, .878,	1.242, 1.569,	2.015)
woba_weights_2017 <- c(0, .877, 1.232, 1.552, 1.980) 
woba_weights_2018 <- c(0, .880, 1.247, 1.578, 2.031)
woba_weights_2019 <- c(0, .870, 1.217, 1.529, 1.940)

#### 2016
model.coef.2016 %<>%
  mutate(expected_woba = (outs_new.prob * woba_weights_2016[1]) + (single_new.prob * woba_weights_2016[2]) + (double_new.prob * woba_weights_2016[3]) + 
           (triple_new.prob * woba_weights_2016[4]) + (home_run_new.prob * woba_weights_2016[5]))
           
model.coef.2016$percentile <- 100 * (scales::rescale(model.coef.2016$expected_woba, to=c(0,1)))
model.coef.2016$qcp <- psych::rescale(model.coef.2016$percentile, mean = 100, sd = 30, df=F)

### 2017
model.coef.2017 %<>%
  mutate(expected_woba = (outs_new.prob * woba_weights_2017[1]) + (single_new.prob * woba_weights_2017[2]) + (double_new.prob * woba_weights_2017[3]) + 
           (triple_new.prob * woba_weights_2017[4]) + (home_run_new.prob * woba_weights_2017[5]))
           
model.coef.2017$percentile <- 100 * (scales::rescale(model.coef.2017$expected_woba, to=c(0,1)))
model.coef.2017$qcp <- psych::rescale(model.coef.2017$percentile, mean = 100, sd = 30, df=F)

### 2018
model.coef.2018 %<>%
  mutate(expected_woba = (outs_new.prob * woba_weights_2018[1]) + (single_new.prob * woba_weights_2018[2]) + (double_new.prob * woba_weights_2018[3]) + 
           (triple_new.prob * woba_weights_2018[4]) + (home_run_new.prob * woba_weights_2018[5]))
           

model.coef.2018$percentile <- 100 * (scales::rescale(model.coef.2018$expected_woba, to=c(0,1)))
model.coef.2018$qcp <- psych::rescale(model.coef.2018$percentile, mean = 100, sd = 30, df=F)

### 2019
model.coef.2019 %<>%
  mutate(expected_woba = (outs_new.prob * woba_weights_2019[1]) + (single_new.prob * woba_weights_2019[2]) + (double_new.prob * woba_weights_2019[3]) + 
           (triple_new.prob * woba_weights_2019[4]) + (home_run_new.prob * woba_weights_2019[5]))
           
model.coef.2019$percentile <- 100 * (scales::rescale(model.coef.2019$expected_woba, to=c(0,1)))
model.coef.2019$qcp <- psych::rescale(model.coef.2019$percentile, mean = 100, sd = 30, df=F)
```

```{r}
batted_ball_results_2015 <- model.coef.2015
batted_ball_results_2016 <- model.coef.2016
batted_ball_results_2017 <- model.coef.2017
batted_ball_results_2018 <- model.coef.2018
batted_ball_results_2019 <- model.coef.2019
```


