---
title: "Contextualizing QC+"
output: html_notebook
---

```{r}
all_results <- rbind(batted_ball_results_2015, batted_ball_results_2016, batted_ball_results_2017,batted_ball_results_2018, batted_ball_results_2019)
```

**Best batter seasons by QC+**
```{r message=FALSE, warning=FALSE}
all_results %>%
  group_by(game_year, player_name) %>%
  summarise(bbe = n(),
            avg_qcp = mean(qcp)) %>%
  filter(bbe > 150) %>%
  arrange(desc(avg_qcp))
```

**Worst batter seasons by QC+**
```{r message=FALSE, warning=FALSE}
all_results %>%
  group_by(game_year, player_name) %>%
  summarise(bbe = n(),
            avg_qcp = mean(qcp)) %>%
  filter(bbe > 150) %>%
  arrange((avg_qcp))
```

**Best pitcher seasons by QC+**
```{r message=FALSE, warning=FALSE}
all_results %>%
  group_by(game_year, pitcher_name) %>%
  summarise(bbe = n(),
            avg_qcp = mean(qcp)) %>%
  filter(bbe > 150) %>%
  arrange(avg_qcp)
```

**Worst pitcher seasons by QC+**
```{r message=FALSE, warning=FALSE}
all_results %>%
  group_by(game_year, pitcher_name) %>%
  summarise(bbe = n(),
            avg_qcp = mean(qcp)) %>%
  filter(bbe > 150) %>%
  arrange(desc(avg_qcp))
```

**Bested batted from each season**
```{r}
for(i in 2015:2019) {
  best_hit <- all_results %>%
    filter(game_year == i) %>%
    dplyr::select(c(player_name, pitcher_name, qcp, game_date)) %>%
    arrange(desc(qcp)) %>%
    slice_head(n = 1)  
  print(best_hit)
  }
```

Video of Donaldson's hit: https://baseballsavant.mlb.com/sporty-videos?playId=4ceff1ef-1e40-4e49-b582-4eaad6b7006e

Video of Cordero's hit: https://baseballsavant.mlb.com/sporty-videos?playId=77b2fd12-d05a-493b-8954-a6e612e93b2c

Video of Sano's hit: https://baseballsavant.mlb.com/sporty-videos?playId=13ac292c-026f-4cb3-999a-a5126148356a

**Worst batted from each season**
```{r}
for(i in 2015:2019) {
  best_hit <- all_results %>%
    filter(game_year == i) %>%
    dplyr::select(c(player_name, pitcher_name, qcp, game_date)) %>%
    arrange((qcp)) %>%
    slice_head(n = 1)  
  print(best_hit)
  }
```

Video of Chapman's hit: https://baseballsavant.mlb.com/sporty-videos?playId=a6d2491d-7554-4b11-8384-9889a781d31b

Video of Hamilton's hit: https://baseballsavant.mlb.com/sporty-videos?playId=1d5503f2-9c8e-4931-80d0-2a5cd890e26e

Video of Kendrick's hit: https://baseballsavant.mlb.com/sporty-videos?playId=429238c5-6d56-47c9-84a6-d3c589ceaaa8


**Best ball by QC+ that was an out**
```{r}
for(i in 2015:2019) {
  best_qcp_out <- all_results %>%
    filter(game_year == i) %>%
    filter(outcome_ec == 0) %>%
    dplyr::select(c(player_name, pitcher_name, qcp, game_date)) %>%
    arrange(desc(qcp)) %>%
    slice_head(n = 1)  
  print(best_qcp_out)
  }
```

Video of Lindor's out: https://baseballsavant.mlb.com/sporty-videos?playId=2f4d7da4-0861-4caa-9310-4476cd18389e - you can see the wind (an unaccounted variable) affecting the ball!

Video of Brantley's out: https://baseballsavant.mlb.com/sporty-videos?playId=84def108-5b88-47e2-91e2-8ba848d0ba09 - the wind also had a major impact here. The announcers thought it was a no-doubter

Video of Garcia's out: https://baseballsavant.mlb.com/sporty-videos?playId=d7694a58-6658-4266-aaf3-3ea213b611de


**Worst ball by QC+ that was a home run**
```{r}
for(i in 2017:2019) {
  best_qcp_out <- all_results %>%
    filter(game_year == i) %>%
    filter(outcome_ec == 4) %>%
    dplyr::select(c(player_name, pitcher_name, qcp, game_date)) %>%
    arrange((qcp)) %>%
    slice_head(n = 1)  
  print(best_qcp_out)
  }
```

Video of Amarista's home run: https://baseballsavant.mlb.com/sporty-videos?playId=71b57e87-c1dc-4b98-922f-ec33ba083b60

Video of Wieters' home run: https://baseballsavant.mlb.com/sporty-videos?playId=c01ba1f6-d3ec-42e4-9c37-b69a4a7780c4

Video of Choo's home run: https://baseballsavant.mlb.com/sporty-videos?playId=63d09954-439f-4698-b308-10eb9df0acb2


**Average QC+ by venue and batter handedness**
```{r message=FALSE, warning=FALSE}
all_results %>%
  group_by(venue_batter.stand) %>%
  summarise(bbe = n(),
            avg_qcp = mean(qcp)) %>%
  arrange(desc(avg_qcp))
```

